---
title: PRISMAstatement README
output:
  md_document:
    variant: gfm
date: "`r format(Sys.time(), '%B %d, %Y')`"
author: Jack O. Wasey
always_allow_html: yes
---
<!-- README.md is generated from README.Rmd. Please edit that file, then:
rmarkdown::render("README.Rmd")
-->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# PRISMAstatement

[![CRAN status](https://www.r-pkg.org/badges/version/PRISMAstatement)](https://cran.r-project.org/package=PRISMAstatement)
[![Travis build status](https://travis-ci.org/jackwasey/PRISMAstatement.svg?branch=master)](https://travis-ci.org/jackwasey/PRISMAstatement)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/jackwasey/PRISMAstatement?branch=master&svg=true)](https://ci.appveyor.com/project/jackwasey/PRISMAstatement)
[![Coverage status](https://codecov.io/gh/jackwasey/PRISMAstatement/branch/master/graph/badge.svg)](https://codecov.io/github/jackwasey/PRISMAstatement?branch=master)
[![lifecycle](https://img.shields.io/badge/lifecycle-stable-blue.svg)](https://www.tidyverse.org/lifecycle/#stable)

PRISMAstatement is a simple package which does one thing: generate publication quality flow charts of study inclusions and exclusions following recommendations of the PRISMA statement. See the vignettes (`vignette("PRISMA")`, `vignette("exclusionflowcharts")`) for more details.

## Installation

You can install the released version of PRISMAstatement from [CRAN](https://CRAN.R-project.org/package=PRISMAstatement) with:

```{r eval=FALSE}
install.packages("PRISMAstatement")
```

## Example

```{r buildexample}
library(PRISMAstatement)
prisma(found = 750,
       found_other = 123,
       no_dupes = 776, 
       screened = 776, 
       screen_exclusions = 13, 
       full_text = 763,
       full_text_exclusions = 17, 
       qualitative = 746, 
       quantitative = 319)
```
