---
title: "Make PRISMA flow chart"
author: "Jack O. Wasey"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: PRISMA.bib
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(DiagrammeR))
```

# The PRISMA Statement

This package plots a [PRISMA](http://prisma-statement.org/) flow chart
    describing the identification, screening, eligibility and inclusion or studies in
    systematic reviews. PRISMA is an evidence-based minimum set of items for reporting
    in systematic reviews and meta-analyses. PRISMA focuses on the reporting of reviews
    evaluating randomized trials, but can also be used as a basis for reporting
    systematic reviews of other types of research, particularly evaluations of
    interventions. [@moher_preferred_2009]

# Generate a PRISMA flow chart

```{r prisma}
library(PRISMAstatement)

prisma(found = 750,
       found_other = 123,
       no_dupes = 776, 
       screened = 776, 
       screen_exclusions = 13, 
       full_text = 763,
       full_text_exclusions = 17, 
       qualitative = 746, 
       quantitative = 319,
       width = 800, height = 800)
```

# References
